\documentclass[../../document.tex]{subfiles}

\begin{document}
    \section{Guided Extraction} \label{sec:extraction:guided}
    We implemented the extraction of \abrv{lcfrs} supertags as described in the previous section and published results in parsing with them \citep{RupMoe21}.
    During the evaluation, we noted that the extracted sets of supertags were rather large and their prediction less accurate than other publications involved with supertagging.
    Following up, we published a generalization of the extraction procedure with the aim of tackling these issues \citep{Rup22}.
    Specifically, the generalization deals with the following two limitations:
    \begin{itemize}
        \item
            Constructing lexical \abrv{lcfrs} rules picks a sentence position for each inner node of the constituent structure according to a fixed strategy (cf.\@ \sref{enum:lcfrs:step3} in \cref{sec:extraction:lcfrs}).
            Such a strategy is now formalized by a \emph{guide} that maps each inner node position of a constituent structure to sentence position in its yield.
            This section introduces multiple strategies to define guides for a given constituent structure; these strategies are called \emph{guide constructors}.
        \item
            LCFRS rules are constructed with constituent symbols as nonterminals, which are then supplemented with annotations during the lexicalization process.
            This section decouples the nonterminals from the other extraction processes and introduces multiple strategies to define them, called \emph{nonterminal constructors}.
    \end{itemize}
    The nonterminal and guide constructors are chosen before the extraction process with a treebank; both choices, as well as a chosen rank transformation, act as parameters for the extraction process across the whole treebank.
    Note that there are certain restrictions to the configuration of the parameters, as explained in \cref{sec:viable-paramters}.
    This section picks up on the published generalizations and describes the extraction procedure in a more formal setting with \abrv{lcfrs}/\abrv{dcp} hybrid grammar rules.
    Similar to the previous section, this extraction process is characterized by four consecutive steps that are applied to each individual constituent tree in the treebank:
    \begin{enumerate}
        \item\label{extraction:hg:ranktrans}
            The constituent tree is processed the chosen rank transformation.
        \item\label{extraction:hg:guide}
            The chosen guide constructor defines a guide for the constituent structure.
        \item\label{extraction:hg:alg}
            An admissible derivation of lexical hybrid grammar rules is constructed using the constituent structure, the guide from the previous step, and the chosen nonterminal constructor.
            The \abrv{lhs} nonterminal at the root of the derivation is collected as an initial nonterminal.
        \item
            In the final step, the rules are collected as a sequence in the order of their lexical symbols.
            All lexical symbols are replaced by a wildcard symbol ``\wildcard{}''.
            In case we aim to extract \abrv{dcp} supertag blueprints, we replace each hybrid grammar rule of the form \(A \to c_1\,c_2\,(\vec{B})\) with the \abrv{dcp} supertag \((A \to c_2 (\vec{B}), |c_1|))\) that does only contain the \abrv{dcp} composition but keeps the length of the \abrv{lcfrs} composition as fanout restriction.
    \end{enumerate}
    The rank transformations for \sref{extraction:hg:ranktrans} are covered in \cref{sec:binarization}.
    \Cref{sec:guides} introduces the guide constructors used in \sref{extraction:hg:guide} and explains which rank transformations are allowed in conjunction with each guide constructor.
    The concept of nonterminal constructors and some examples are discussed in \cref{sec:ntconstructors}; the remainder of \sref{extraction:hg:alg} is explained in \cref{sec:extractionalg}.
    

    \subfile{41-guides.tex}
    \subfile{42-nonterminals.tex}
    \subfile{43-extractionalg.tex}

    \ifSubfilesClassLoaded{%
        \printindex
        \bibliography{../references}%
    }{}
\end{document}