\documentclass[../../document.tex]{subfiles}

\begin{document}
    \section{Extracting and Lexicalizing Simple \abrv{Lcfrs}} \label{sec:extraction:readoff}
    This extraction scheme for supertags with \abrv{lcfrs} rules was the product of a joint work with Richard Mörbitz. \citep{MoeRup20,RupMoe21}
    It is based on simple \abrv{lcfrs} read-off from treebanks as shown by \citet{KalMai13} followed by a per-tree lexicalization inspired by the procedures presented by \citet{EngMalMan18}.

    Here, an \abrv{lcfrs} supertag is a tuple \((r, s)\) consisting of a lexical rule \(r\) and a transportation marker \(s\).
    The lexical symbol in \(r\) is a sentence position in the yield of the constituent tree.
    The boolean transportation marker \(s\) is used to undo the lexicalization procedure and obtain a constituent tree from a derivation of supertags.
    We describe the extraction procedure independently for each constituent tree in a treebank in six steps:\footnote{
        The original formulation contained an additional step (that was called \emph{dechain}) to remove unary non-lexical rules.
        Our formulation of binarization already merges unary nodes in the same way as this step, and we can just skip it with equal results.
    }
    \begin{enumerate}
        \item The constituent tree is binarized.
        \item
            A simple (binary) \abrv{lcfrs} derivation is read from the constituent tree.
            The derivation is instantiated in the sequence of tokens of the constituent tree.
            For a sentence of length \(n\), it contains exactly \(n-1\) inner nodes with non-lexical rules and \(n\) leaves with lexical rules.
        \item\label{enum:lcfrs:step1}
            The derivation's leaves are dismissed and their lexical symbols are moved into their parents by each replacing a variable.
            The \abrv{rhs} nonterminals of all rules are adjusted to accommodate the removed rules.
        \item\label{enum:lcfrs:step2}
            The derivation is then transformed by \emph{transporting} the lexical symbols into remaining unlexical rules according to a fixed strategy.
            During this step, the rules' \abrv{rhs} nonterminals are annotated according to the transportation.
            Moreover, each node with rule \(r\) is replaced by a tuple \((r, s)\) where \(s\) is true if and only if the node was equipped another lexical symbol before the step.
        \item\label{enum:lcfrs:step3}
            One rule in the derivation is equipped with two lexical symbols.
            It is split by replacing one of the symbols with a variable and adding a new nullary rule.
        \item
            In the final step, the rules are collected as a sequence in order of the positions according to the instantiation.
    \end{enumerate}
    In the following, the steps \crefrange{enum:lcfrs:step1}{enum:lcfrs:step3} are described in more detail.


    \subsection{Lexicalizing \abrv{Lcfrs} Derivations}\label{sec:extraction:readoff:lex}
    The lexicalization scheme is based on \citet{MoeRup20}.
    However, we ignore all weights and perform lexicalization on individual derivations rather than on a grammar induced from the entire treebank.
    More specifically, we directly read off a set of lexical rules from each tree in the treebank; then the union of these rules forms our uni-lexical LCFRS $\lexG$.
    In contrast to that, \citet{MoeRup20} first induce an LCFRS $G$ from the entire treebank and then lexicalize $G$.
    Thus $\lexG$ may have a different language than the lexicalization of $G$.

    We obtain a set of uni-lexical rules from each tree $t$ in the treebank by the following procedure.
    \begin{enumerate}
        \item Binarize the tree.
        The symbol $|$ is appended to constituents that result from binarization (this reflects Markovization with a vertical context of $1$ and a horizontal context of $0$).
        \label[step]{step:binarize}
        \item Transform the tree into an LCFRS derivation using the standard technique for induction of LCFRS \cite{MaierSogaard08}.
        \label[step]{step:induce}
        \item Collapse every chain of monic rules; the nonterminals of each chain are combined to a new nonterminal.
        \label[step]{step:dechain}
        \item Remove every terminating rule that has a parent and insert the terminal from its composition into the parent.%
        \label[step]{step:fuseterm}%
        \item Propagate terminals from double-lexical terminating rules into non-lexical branching rules.
        All rules in the resulting derivation are lexical.
        \label[step]{step:propterm}
        \item Split all remaining double-lexical terminating rules into two uni-lexical rules.
        All rules in the resulting derivation are uni-lexical.
        The resulting derivation is called $d_{\mathrm{lex}}(t)$.
        \label[step]{step:split}
        \item Read off the rules of $d_{\mathrm{lex}}(t)$; call them $R(t)$.
        \label[step]{step:readoff}
    \end{enumerate}

    These steps are defined such that in the LCFRS formed by $R(t)$, $d_{\mathrm{lex}}(t)$ is a derivation for the sentence of $t$.
    Moreover, we are able to reconstruct $t$ from $d_{\mathrm{lex}}(t)$ by reverting \crefrange{step:split}{step:binarize} (we will give the details later).

    Finally, we obtain the uni-lexical LCFRS $\lexG$ by combining the rules $R(t)$ for each tree $t$.
    The initial nonterminals of $\lexG$ are all left-hand sides of roots of $d_{\mathrm{lex}}(t)$.

    Let $t$ be a tree in the treebank.
    \Cref{step:binarize,step:induce} and their reversal are standard techniques for trees and LCFRS.
    After applying them to $t$, we obtain an LCFRS derivation in which each occurring rule is either of the form
    \begin{itemize}[nosep]
        \item \(A \to (\sigma)\), where $\sigma$ is a terminal and \(A\) is the part-of-speech tag of $\sigma$,
        \item \(A \to c(B_1)\) where \(\fanout(A) = \fanout(B)\) and \(c = \id_{\fanout(A)}\), or
        \item \(A \to c (B_1, B_2)\) where \(c\) contains no terminals and none of \(B_1, B_2\) is an initial nonterminal.
    \end{itemize}
    Let us denote this derivation by $d$.

    In the following, we describe \crefrange{step:dechain}{step:split} of the above procedure in more detail (showing examples in \crefrange{fig:dechain}{fig:split}) and also glimpse at how the individual steps are reverted.


    \paragraph{\Cref{step:dechain}.}
    We \emph{repeatedly} replace parts in \(d\) of the form
    \(A \to c (B) \big( B \to c' \big) \) by \( A\text+B \to c' \), and
    \(A \to c (B) \Big( B \to c'(C_1, C_2) \big( \ldots \big) \Big) \) by \( A\text+B \to c' (C_1, C_2) \big( \ldots \big) \),
    until there is no monic rule in \(d\) left.
    If the occurrence of \(A \to c (B)\) is not the root of \(d\), then the corresponding nonterminal in the parent's rhs is replaced by \(A\text+B\).\footnote{
        Note that root nodes in the derivation may be collapsed, this is why we use LCFRS with \emph{multiple} initial nonterminals.
    }
    After this step, there are only branching rules and terminating rules in \(d\).
    \Cref{fig:dechain} shows an example for this step.

    This step is easy to reverse, as the composition of every removed rule is $c = \id_{\fanout(B)}$.
    We give the formal description in \cref{unlex:dechain}.

    \paragraph{\Cref{step:fuseterm}.}
    We remove every non-root occurrence $r$ of a terminating rule $A \to (\sigma)$ in $d$.
    Let $r$ be the $i$th child of its parent (with $i \in [2]$), then we replace the parent's composition $c$ by $\sem{c}_i(\sigma)$ and remove the $i$th nonterminal in the parent's rhs.

    %We replace all occurrences in \(d\) of the form
    %\begin{itemize}
    %  \item \(A \to c(B_1, B_2) \Big(B_1 \to (\sigma_1), B_2 \to (\sigma_2)\Big)\) by \(A \to \sem{c}(\sigma_1, \sigma_2)\),
    %  \item \(A \to c(B_1, B_2) \Big(B_1 \to (\sigma_1), \ldots \Big)\) by \(A \to \sem{c}_1(\sigma_1)\,(B_2)\Big( \ldots \Big)\), and
    %  \item \(A \to c(B_1, B_2) \Big(\ldots, B_2 \to (\sigma_2) \Big)\) by \(A \to \sem{c}_2(\sigma_2)\,(B_1)\Big( \ldots \Big)\).
    %\end{itemize}
    We note that the parent becomes lexical, and after this step, every rule in $d$ is either branching or lexical.
    Moreover, every terminal rule in $d$ is either double-lexical (if both children were removed) or the root of \(d\) (and thus its only node).
    \Cref{fig:fuseterm} shows an example for this step.

    Clearly, this step loses information, namely the left-hand sides of the removed rules.
    These nonterminals are part-of-speech tags (that may be enriched with nonterminals of collapsed monic rules from the previous step).
    For the reversal of this step, we opted to predict them along with the supertags as part of the supertagger.
    The formal description of the reversal is given in \cref{unlex:fuseterm}.

    \paragraph{\Cref{step:propterm}.}
    For each occurrence \(r\) of a branching rule $A \to c(A_1, A_2)$ in \(d\), let us consider the occurrence \(t\) of the leftmost terminating rule (i.e.\@ \(t\) is a leaf) that is reachable via the second successor of \(r\).
    For example, in \cref{fig:propterm:pre}, the two binary rules (\(r\)) are end points of gray arrows; these arrows start at the mentioned leaves (\(t\)).
    Our goal is to remove one terminal from $t$ and propagate it all the way up to $r$.
    For this, at each node \(s\) on the path from \(t\) to \(r\) (from bottom up):
    \begin{itemize}[nosep]
        \item If \(s\) is \(t\), we remove the leftmost terminal in the rule's composition at \(s\).
        \item
        If \(s\) is neither \(t\) nor \(r\), we insert the last removed terminal right before the variable \(\x_1\) and then remove the leftmost terminal in the rule's composition at \(s\).

        We note that if the rule at $s$ is monic and the variable $\x_1$ occurs right of the terminal in its composition, then we propagate a different terminal than the one received from the child.
        In order to be able to reverse this step, we need to remember whether the terminal in the rule's composition stayed the same or was swapped with the terminal received from the child.
        In the following, we consider this information as part of the rule (cf.\@ the gray annotation \textsuperscript{swapped} in \cref{fig:propterm}).
        \item If \(s\) is \(r\), we insert the last removed terminal right before the variable \(\y_1\) in the rule's composition at \(s\).
    \end{itemize}
    If \(s \neq r\), let \(s'\) be the parent of $s$ and $s$ the \(i\)th child of $s'$.
    If, after removal of a terminal at \(s\), the first component in the composition is empty:
    \begin{itemize}[noitemsep,topsep=1pt]
        \item we annotate the lhs nonterminal at \(s\) and the \(i\)th rhs nonterminal at \(s'\) with $^-$ and remove the empty component, and
        \item if \(i = 1\) (resp.\@ \(i = 2\)), we remove \(\x_1\) (resp.\@ \(\y_1\)) and replace every other occurrence of \(\x_i\) by \(\x_{i-1}\) (resp.\@ \(\y_j\) by \(\y_{j-1}\)) at $s'$.
    \end{itemize}
    Otherwise, we annotate the nonterminals with \(^+\).

    We note that the rule at $r$ is uni-lexical and branching now, the rule at \(t\) is uni-lexical and terminating, and the number of terminals in each rule between them did not change.
    After this step, every rule in $d$ is lexical.
    \Cref{fig:propterm} shows an example for this step.

    There is a suitable leaf $t$ for every branching rule $r$.
    Intuitively, this holds since (a) after \cref{step:fuseterm} every leaf of $d$ is a double-lexical rule and (b) for each branching rule we first go to its second successor and then follow the path of first successors until we reach a leaf.
    Here, (a) guarantees that there exists a double-lexical rule for each branching rule and (b) guarantees that a terminal in each double-lexical rule is transported to at most one branching rule, thus at most one terminal is removed from it.
    We refer the more interested reader to consult the proof of correctness by \citet{EngMalMan18};
    this proof also applies to our method.

    The reversal of this step removes all annotation ($^+$, $^-$, and \textsuperscript{swapped}) and restores each composition in $d$ to its original form.
    We note that the original composition can be obtained deterministically; the construction is given in \cref{unlex:propterm}.

    \paragraph{\Cref{step:split}.}
    We replace the rightmost terminal $\sigma_2$ in the composition of each double-lexical terminating rule by a variable and add a new nonterminal $A^{\text{R}}$ to the rule's right-hand side (making it a uni-lexical monic rule).
    Then we insert the rule $A^{\text{R}} \to (\sigma_2)$ as a child.
    %Each occurrence of the form
    %  \(A \to (\sigma_1 \sigma_2)\) is replaced by \(A \to (\sigma_1 \x_1) (A^\text{R}) \Big(A^\text{R} \to (\sigma_2)\Big)\), and
    %  \(A \to (\sigma_1, \sigma_2)\) is replaced by \(A \to (\sigma_1, \x_1) (A^\text{R}) \Big(A^\text{R} \to (\sigma_2)\Big)\),
    %where \(A^\text{R}\) is a new nonterminal.
    After this step, every rule in $d$ is uni-lexical.
    \Cref{fig:propterm:post,fig:split} show an example for this step.
    The reversal of this step is straightforward.
    %We formally describe it in \cref{unlex:split}.
\end{document}